%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the LaTeX source for
% "A 2D Nearest-Neighbor Quantum Architecture for Factoring"
% submitted to the Reversible Computing Workshop (RC 2012)
% based on Spring-Verlag's Lecture Notes in Computer Sciences template
% typeinst.tex
% Paul Pham and Krysta Svore
% 14 March 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[runningheads]{llncs}
% Suppress page numbers
%\documentclass[a4paper]{llncs}
% For arXiv, and eprint support
\documentclass{article}

\usepackage{amssymb}
\usepackage{amsthm}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{eprint}
%\usepackage[osf]{mathpazo} % Use Palatino / Euler fonts
%\usepackage{multirow}
%\usepackage[retainorgcmds]{IEEEtrantools}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.beyer, christine.guenther,|
\urldef{\mailsb}\path|frank.holzwarth, anna.kramer, erika.siebert-cole, lncs}@springer.com|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

% Right brace for multirows in tables / arrays
\newcommand\coolrightbrace[2]{%
\left.\vphantom{\begin{matrix} #1 \end{matrix}}\right\}#2}

% To fix Qcircuit target with new Xypic
\newcommand{\targfix}{\qw {\xy {<0em,0em> \ar @{ - } +<.39em,0em>
\ar @{ - } -<.39em,0em> \ar @{ - } +
<0em,.39em> \ar @{ - }
-<0em,.39em>},<0em,0em>*{\rule{.01em}{.01em}}*+<.8em>\frm{o}
\endxy}}

% To get Roman uppercase Greek characters
\newcommand{\X}[1]{$#1$}

\input{Qcircuit}

\newcommand{\braket}[2]{\langle #1|#2 \rangle}
\newcommand{\normtwo}{\frac{1}{\sqrt{2}}}
\newcommand{\norm}[1]{\parallel #1 \parallel}
\newcommand{\email}[1]{\href{mailto:#1}{#1}}
\theoremstyle{plain} \newtheorem{lemma}{Lemma}

\begin{document}

%\mainmatter  % start of an individual contribution

% first the title is needed
\title{A 2D Nearest-Neighbor Quantum Architecture for Factoring}

% a short form should be given in case it is too long for the running head
%\titlerunning{A 2D Nearest-Neighbor Quantum Architecture for Factoring}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Paul Pham\\
University of Washington\\
Quantum Theory Group\\
Box 352350, Seattle, WA 98195, USA,\\
\email{ppham@cs.washington.edu},\\
\url{http://www.cs.washington.edu/homes/ppham/}
\and
Krysta M. Svore\\
Microsoft Research\\
Quantum Architectures and Computation Group\\
One Microsoft Way, Redmond, WA 98052, USA\\
\email{ksvore@microsoft.com},\\
\url{http://research.microsoft.com/en-us/people/ksvore/}
}
% if the list of authors exceeds the space for a headline,
% an abbreviated author list is needed
%\authorrunning{P. Pham \and K.M. Svore}
% (feature abused for this document to repeat the title also on left hand pages)

\maketitle

This document responds to comments by Referee 1, which were received on
November 30, 2012. These comments are quoted and responded to below.  We thank the Referee for the constructive feedback and suggestions.

\section{General Comments}

\begin{quote}
In this paper, the authors present in detail a new version of the
modular exponentiation component of Shor's algorithm, with attention
to the constraints of a 2-D planar graph of moderate degree for qubit
connectivity.  They use teleportation-based fanout to move qubits
around within the machine.  The modular addition uses a novel method,
depending upon carry-save arithmetic and a small triangular lattice as
its unit cell.

Perhaps the most novel part of the arithmetic is the approach to
modular multiplication.  In a traditional modular multiplication of
two numbers $x$ and $y$, the $n^2$ bit-wise products $x_i y_j$ are calculated
and laid out in a trapezoid shifted to give the correct column (power
of two), such that each entry is $2^i 2^j x_i y_j$, which of course can
be represented by a single bit in the right place.  Then the columns
are added, creating a 2n-bit number, which then must be further
operated on to perform the modulo $N$ operation.  ($N$ is assumed to be an
$n$-bit number with the high-order bit being a one.)

In contrast, in this approach, each of the $n^2$ entries is a full $n$-bit
number, $(2^i 2^j \bmod m) x_i y_j$.  By adding those numbers using their
modular circuit, the full $x y \bmod N$ value can be calculated directly.
As proposed, this requires $O(n^3)$ bits (qubits) in the register.  By
combining the $n^2$ partial results in a log-depth tree structure, the
depth for a modular multiplication becomes $O(\log n)$ times the constant
depth of their 3-to-2 carry-save modular operation.

Up through section 5, I was prepared to recommend nearly immediate
acceptance of the paper.  The writing is clear and elegant and the
technical work both valuable and polished.  In sections 6 and 7,
however, I have some doubts about the technical work, and the writing
seems a bit more rushed, and the tail of the paper is not yet
satisfactory.

Most importantly, it is disappointing that the authors have not
produced a more complete estimate of the number of qubits (resources)
required, as well as the actual circuit depth.  A rough estimate, at
least, should be very achievable given the level of detail already
developed in the paper.  The authors mention this as future work, but
without it, the value of the paper is substantially diminished, and it
does not seem unreasonable to expect it to appear here.

\end{quote}

{\it We thank the referee for the valuable feedback. More detailed numerical
upper bounds for circuit resources are provided throughout the paper for each
building block and the overall architecture. To do so, we have clarified our architectural
model to include modules.}

\section{Other significant technical comments:}

\begin{quote}
The authors appear not to have noticed that half of the $n^2$ numbers in
their multiplier require only a single bit.  As long as $i+j \le n$, the
modulo operation results in the same number, allowing us to avoid
using a full $n$-bit number: $(2^i 2^j \bmod m) x_i y_j = 2^i 2^j x_i y_j$.
\end{quote}

{\it Thank you for pointing this out, it is incorporated into the resource estimates for
the modular multiplier, in the partial product creation section.}

\section{Section 2.2}

\begin{quote}
``this 'consumes' the cat state'' -- this sentence is confusing.
\end{quote}

{\it We meant that the cat state remains entangled with the original source
qubit and its fanned-out, entangled copies. We have since discovered
via personal conversation with Aram Harrow and Dan Browne that this
statement is no longer true. It is possible to ``un-fanout'' and therefore
disentangle the cat state from the source and target qubits, allowing us
to reuse this state.  However, we have not been able to confirm the
exact circuit at the time of this writing, so we have removed this sentence.
Instead, we discuss an alternative method of unfanout, which is alternating
teleports and CNOTs in a logarithmic-depth binary tree.}

\section{Section 3}

\begin{quote}
  Choi and Van Meter were not the first to consider 2-D architectures;
most of the solid-state proposals and even some of the ion trap
proposals worked with a 2-D layout.  Kielpinski's 2002 proposal might
have been the first 2-D architecture.  Working out exact algorithms on
the structures came later, but papers by Kubiatowicz's group and
Chong/Oskin clearly included at least some level of work on the
movement of qubits in a planar system, albeit with less attention to
the abstraction of logical qubit connectivity.
\end{quote}

{\it 
%We've added a reference to the 2002 Nature paper by Kielpinski et al.
%We have also included references to several quantum architecture papers in this section.
We intended to say that Choi and Van Meter were the first to construct an adder explicitly optimized for a 2D architecture of qubits.
We have reworded this paragraph to accurately reflect their contribution and added appropriate references.}

%KRYSTA TODO: Can you add references to quantum architecture papers by
%Kubi/Chong/Oskin on 2D implementations, since you are probably more
%familiar with those works?

\begin{quote}
  I'm not sure a modular adder is "extended" to do modular
exponentiation.  "Composed" or "used", perhaps?
\end{quote}

{\it We have reworded this sentence.}

\begin{quote}
  Your citation of Gossett has no year.
\end{quote}

{\it Corrected.}

\begin{quote}
"all other factoring implementations"?  That's a rather broad
characterization.  Cleve and Watrous long ago proposed using a
parallel reduction tree of multiplications before the QFT.  Van Meter
and Itoh investigated in detail the tradeoffs in resource consumption
for doing this.
\end{quote}

{\it We have clarified this sentence by referring to all nearest-neighbor factoring
implementations. By ``implementations,'' we mean a concrete mapping to an
architecture, such as those given by [Fowler et al. 2004] and [Kutin 2006].
While we do acknowledge the Cleve and Watrous paper, which gives
similar parallel results to those in the Kitaev-Shen-Vyalyi book, both
assume arbitrary connectivity of qubits.}

\begin{quote}
As far as I am aware, no one has worked out the details of Draper's
transform adder taking into account the need to do Solovay-Kitaev
decomposition.  This may add a very large factor to the execution
time.
\end{quote}

{\it Agreed, we are not aware of anyone working out compilation of the Draper
transform adder to a universal gate set, such as the Clifford group plus
$\pi/8$ gates. It's been shown that compilation will require $O(\log (1/\epsilon))$ overhead, where $\epsilon$ is the required precision. 
Throughout, we choose to avoid use of the QFT in our circuit due to the compilation required for each small rotation gate.  We have added a sentence regarding this compilation cost.
%how to compile more efficiently
%than Solovay-Kitaev using more time and space, but it would still not be
%more efficient than the Gossett adder.
}

\begin{quote}
Do you think the Zalka approximate multiplier approach actually works?
\end{quote}

{\it It is likely that something similar to the Zalka approximate
multiplier actually works in practice for the majority of input values,
if not the exact implementation described in
the Zalka and Kutin papers. However, a rigorous theoretical argument, or
empirical verification by simulation, has not yet appeared in the literature.
}

\begin{quote}
While it's okay to include a "forward reference" to a forthcoming
paper of your own that carries more detail, you can't ask us to "refer
to" it!
\end{quote}

{\it Agreed. Citation has been removed.}

\begin{quote}
I don't think the BKP ``exact" circuit guarantees a result, does it?
Shor's own original algorithm only probabilistically gives the correct
answer, and that probability is a matter of some debate in the
literature.  Papers by Fowler (2004), Miquel (1996), Garcia-Mata
(2007) and others provide different estimates.
\end{quote}

{\it Correct.  In this work we only analyze the resource cost for a single round of QPF.  The probability of achieving a correct answer may vary.}

\begin{quote}
The journal style will ultimately dictate this, but when the
bibliographic labels in the text are alphabetic, the references are
usually ordered alphabetically.
\end{quote}

{\it We have updated the style.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Section 4}

\begin{quote}
It would be worth pointing out that qubit 0 is the low-order qubit.
\end{quote}

{\it Corrected.}

\begin{quote}
``At the level of bits, a CSA...'' this sentence is awkward, reword.
\end{quote}

{\it Corrected.}

\begin{quote}
Fig. 4 shows the layout, but it doesn't exactly match the circuit of
Fig. 3.  Having the exact circuit to accompany Fig. 4 would be useful.
\end{quote}

{\it The layout in Fig. 4 matches Fig. 3 if you look at the outgoing qubit values in Fig. 3.  We have clarified this in the text.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Section 5}

\begin{quote}
It feels somewhat like the phrasing on constant depth is a bit
misleading in this section.  Please reread and make sure it is easy
for the reader to follow your claims.
\end{quote}

{\it Can you please provide specific examples of misleading phrasing?}

\begin{quote}
Proof of Lemma 1: "O.ur" typo.
\end{quote}

{\it Corrected.}

\begin{quote}
A little attention to classical versus quantum addends in this section
might help the reader.
\end{quote}

{\it Unfortunately, due to time constraints, we were unable to incorporate this feedback into our revisions.}

\begin{quote}
Fig. 5: labeling the lines themselves as "Layer 1", "Layer 2",
etc. might help the reader.
\end{quote}

{\it This is a great improvement and has been incorporated.}

\begin{quote}
"at no layer generates" --> "no layer generates"?
\end{quote}

{\it Corrected.}

\begin{quote}
Fig. 6: Swap the left and right ends of this figure to make it
correspond to Fig. 5 as closely as possible, and point out this
correspondence by also labeling the "layers" and the time axis here.
"FANOUT RAIL" -- of which variable(s)?
\end{quote}

{\it Unfortunately, due to time constraints, we were not able to incorporate this suggestion.
The fanout rails are of the only qubits connected to them, namely, $v_4$, $u_4$, and $v_5$, respectively.}

\begin{quote}
My estimate here is that spatial resources are 29n qubits, temporal
resources are 12n Toffoli gates.  Is that about right?
\end{quote}

PAUL TODO
{\it Add response.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Section 6}

\begin{quote}
If you are worried that "quantum number" will mislead some readers,
would "quantum integer" be better?
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Bottom of 6.1 is the place to expand the discussion to incorporate the
above on the size of partial products and their impact on resource
utilization.
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Your white tiles are hard to see on my printout.
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Important: Your second and third rules under "black tiles" appear to
conflict.  Reword to be precise.
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Bottom of Sec. 6: more discussion of the size and number of addends
and resource consumption is needed.
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Fig. 10: In earlier figures, arrows were used only to indicate addend
motion via teleportation, correct?  A different symbol to indicate
actual multiplication would make the figure clearer.  Otherwise,
asserting that an arrow itself is "log n depth" will confuse the
reader.  The tree approach used in this figure is not original, and in
fact dates to the 2000 paper by Cleve and Watrous, at least.
\end{quote}

{\it We have clarified the use of arrows in Fig. 10.}

{\it We do not claim that this tree structure is original, merely that it is
now possible to do the necessary communication (teleportation and fan-out)
in constant depth. That is, the tree structure before was a theoretical
construction, and now based on the previous nearest-neighbor implementations
of modular arithmetic, correspond to an actual physical tree structure if
this architecture were to be fabricated and observed to run over time.}

\begin{quote}
Sections 7, 8 and 9 should be extended with more discussion, and
actual resource consumption figures.  Some of the papers cited in
Fig. 11 (which ought to be a table, not a figure) provide detailed
estimates on resources (depth and width), as do Beckman et al. (1996,
uncited here, but should be) and Van Meter and Itoh.  Both of those
latter papers offer several configurations that may make direct
comparison tricky.
\end{quote}

PAUL TODO
{\it Add response.}

\begin{quote}
Overall, as noted, this paper will be a valuable contribution to the
literature once these issues are addressed.
\end{quote}

{\it We thank the referee for the valuable feedback.}

\bibliography{factor2d}
\bibliographystyle{tocplain}

\end{document}
